
```{r}
library(dplyr)
library(tidyverse)
library(DBI)
```

Reading in the data and combining the data frames using the code provided.
```{r}
d1 =read.table("student-mat.csv",sep=";",header=TRUE)
d2 =read.table("student-por.csv",sep=";",header=TRUE)

d3=merge(d1,d2,by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))
print(nrow(d3)) # 382 students
```

Reading in the data and combining using functions from the tidyverse. It gives me a message "Warning: Detected an unexpected many-to-many relationship between `x` and `y`."
```{r}
mat2 <- read.csv("student-mat.csv", sep = ";", header = TRUE)
por2 <- read.csv("student-por.csv", sep = ";", header = TRUE)

combined <- inner_join(mat2, por2, by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))
```

Using an inner_join() on all variables other than G1, G2, G3, paid, and absences. 
```{r}
join_exclusions <- setdiff(colnames(mat2), c("G1", "G2", "G3", "paid", "absences"))
exercise_combine <- inner_join(mat2, por2, by = join_exclusions)
```

Next, for the math data, Portuguese, and combined data, choosing four categorical variables of interest and converting those into factor variables in each tibble using the mutate() function.
```{r}
mat2 <- mat2 |>
  mutate(romantic = as.factor(romantic),
         guardian = as.factor(guardian),
         famsup = as.factor(famsup),
         reason = as.factor(reason))

por2 <- por2 |>
  mutate(romantic = as.factor(romantic),
         guardian = as.factor(guardian),
         famsup = as.factor(famsup),
         reason = as.factor(reason))

exercise_combine <- exercise_combine |>
  mutate(romantic = as.factor(romantic),
         guardian = as.factor(guardian),
         famsup = as.factor(famsup),
         reason = as.factor(reason))
```

Looking at how the data is stored.
```{r}
str(exercise_combine)
```

Documenting missing values in the data.
```{r}
sum_na <- function(column) {
  sum(is.na(column))
}
na_counts <- exercise_combine |>
  summarize(across(everything(), sum_na))
na_counts
```

Creating a one-way contingency table. From this table we see that there are 221 people from the data set who are not in a romantic relationship, and 99 people who are.
```{r}
one_way <- table(exercise_combine$romantic)
one_way
```

Creating a two-way contingency table. From this, we can see that 
```{r}
two_way <- table(exercise_combine$romantic, exercise_combine$reason)
two_way
```
Creating a three-way contigency table. From this we gather that 
```{r}
three_way <- table(exercise_combine$romantic, exercise_combine$reason, exercise_combine$guardian)
three_way
```

Creating a conditional two-way table using table() by subsetting the data by females and then creating the table.
```{r}
sex_filter <- filter(exercise_combine, sex == "F")

female_table <- table(sex_filter$romantic, sex_filter$guardian)
female_table
```

Create a conditional two-way table using table() by creating a three-way table and subsetting it.
```{r}
subset_table <- table(exercise_combine$romantic, exercise_combine$guardian, exercise_combine$reason)

filter_subset_table <- subset_table[,"mother",]
filter_subset_table
```

Creating a two-way contingency table using group_by() and summarize().
```{r}
exercise_combine |>
  group_by(romantic, reason) |>
  drop_na(romantic, reason) |>
  summarize(count = n()) |>
  pivot_wider(names_from = reason, values_from = count)
```
Create stacked bar graph.
```{r}
stacked <- ggplot(data = exercise_combine |> drop_na(sex, famsup), aes(x = sex, fill = famsup)) 
stacked + geom_bar() + labs(title = "Stacked Bar Graph of Family Educational Support by Sex", x = "Sex", y = "Family Educational Support")
```

Create side-by-side bar graph
```{r}
side <- ggplot(data = exercise_combine |> drop_na(school, reason), aes(x = school, fill = reason)) + geom_bar(position = "dodge") + labs( title = "Side-By-Side Bar Graph of Reason by School", x = "School", y = "Reason") + scale_fill_discrete("Reason")
side
```

# Numeric Variables (and across groups)
```{r}
exercise_combine |>
  summarize(mean_age = mean(age),
            median_age = median(age),
            mean_absences = mean(absences.x),
            median_absences = median(absences.x),
            mean_G1 = mean(G1.x),
            median_G1 = median(G1.x))
```

```{r}
exercise_combine |>
  filter(sex == "F") |>
   summarize(mean_age = mean(age),
            median_age = median(age),
            mean_absences = mean(absences.x),
            median_absences = median(absences.x),
            mean_G1 = mean(G1.x),
            median_G1 = median(G1.x))
  
```
```{r}
exercise_combine |>
  group_by(sex) |>
   summarize(mean_age = mean(age),
            median_age = median(age),
            mean_absences = mean(absences.x),
            median_absences = median(absences.x),
            mean_G1 = mean(G3.x),
            median_G1 = median(G3.x))
```

```{r}
exercise_combine |>
  group_by(sex,school) |>
   summarize(mean_age = mean(age),
            median_age = median(age),
            mean_absences = mean(absences.x),
            median_absences = median(absences.x),
            mean_G1 = mean(G3.x),
            median_G1 = median(G3.x))
```

```{r}
numeric_variables <- exercise_combine[, sapply(exercise_combine, is.numeric)]
numeric_variables
```

```{r}
correlation_matrix <- cor(numeric_variables)
correlation_matrix
```

Histogram of Age by Sex
```{r}
histogram_age <- ggplot(exercise_combine, aes(x = age)) + geom_histogram(aes(fill = sex)) + labs(title = "Histogram of Age by Sex")
histogram_age
```

Kernel density plot
```{r}
kernel_age <- ggplot(exercise_combine, aes(x = age)) + geom_density(alpha = 0.5, aes(fill = sex)) + labs(title = "Kernel Density Plot of Age by Sex")
kernel_age
```

Age boxplot
```{r}
boxplot_age <- ggplot(exercise_combine, aes(x = sex, y = age, fill = sex)) + geom_boxplot() + labs(title = "Boxplot of Age by Sex", x = "Sex", y = "Age")
boxplot_age
```
Histogram of
```{r}
histogram_absences <- ggplot(exercise_combine, aes(x = absences.x)) + geom_histogram(aes(fill = sex)) + labs(title = "Histogram of Absences by Sex")
histogram_absences
```

Kernel Density Plot of 
```{r}
kernel_absences <- ggplot(exercise_combine, aes(x = absences.x)) + geom_density(alpha = 0.5, aes(fill = sex)) + labs(title = "Kernel Density Plot of Absences by Sex")
kernel_absences
```

Boxplot of 
```{r}
boxplot_absences <- ggplot(exercise_combine, aes(x = sex, y = absences.x, fill = sex)) + geom_boxplot() + labs(title = "Boxplot of Absences by Sex", x = "Sex", y = "Absences")
boxplot_absences
```

Scatterplot relating G3 variable to other numeric variables
```{r}
scatterplot_G3_age <- ggplot(exercise_combine, aes(x = age, y = G3.x, color = sex)) + geom_point() + geom_jitter(width = 0.1, alpha = 0.3) + labs(title = "Scatterplot of G3 and Age by Sex", x = "Age", y = "G3")
scatterplot_G3_age
```

```{r}
scatterplot_G3_failures <- ggplot(exercise_combine, aes(x = failures, y = G3.x, color = school)) + geom_point() + geom_jitter(width = 0.1, alpha = 0.3) + labs(title = "Scatterplot of G3 and Failures by School", x = "Failures", y = "G3")
scatterplot_G3_failures
```

```{r}
G3_age_facet <- ggplot(exercise_combine, aes(x = age, y = G3.x, color = sex)) + geom_point() + geom_jitter(width = 0.1, alpha = 0.3) + labs(title = "Scatterplot of G3 and Age by Sex", x = "Age", y = "G3") +
  facet_wrap(~ internet)
G3_age_facet
```

```{r}
G3_failures_facet <- ggplot(exercise_combine, aes(x = failures, y = G3.x, color = school)) + geom_point() + geom_jitter(width = 0.1, alpha = 0.3) + labs(title = "Scatterplot of G3 and Failures by School", x = "Failures", y = "G3") +
  facet_wrap(~ famsize)
G3_failures_facet
```

```{r}
G3_age_facet2 <- ggplot(exercise_combine, aes(x = age, y = G3.x, color = sex)) + geom_point() + geom_jitter(width = 0.1, alpha = 0.3) + labs(title = "Scatterplot of G3 and Age by Sex", x = "Age", y = "G3") +
  facet_grid(rows = vars(internet), cols = vars(school))
G3_age_facet2
```

```{r}
G3_failures_facet2 <- ggplot(exercise_combine, aes(x = failures, y = G3.x, color = school)) + geom_point() + geom_jitter(width = 0.1, alpha = 0.3) + labs(title = "Scatterplot of G3 and Failures by School", x = "Failures", y = "G3") +
  facet_grid(rows = vars(famsize), cols = vars(activities))
G3_failures_facet2
```

